---
title: "data_wrangling"
format: html
---

```{r}
library(tidyverse)
library(here)
library(janitor)

host_cities   <- read_csv(here("inst/extdata/eurovision_host_cities.csv"))
participants  <- read_csv(
  here("inst/extdata/eurovision_participants.csv"),
  show_col_types = FALSE,
  col_types       = cols(.default = col_character())
)
results       <- read_csv(here("inst/extdata/eurovision_results.csv"))
broadcasts    <- read_csv(here("inst/extdata/eurovision_broadcasts.csv"))
```

# Inspect Host Cities df

Manual inspection and edits

```{r}
host_cities <- host_cities %>% 
  mutate(
    venue = case_when(
      year == 1965 ~ "Sala di Concerto della RAI",
      .default = venue
    ),
    host_city = case_when(
      year == 1965 ~ "Naples, Italy",
      year == 1974 ~ "Brighton, United Kingdom",
      year == 2010 ~ "Oslo, Norway",
      year == 2013 ~ "Malmö, Sweden",
      .default = host_city
    )
  )

# rewrite over 
write_csv(host_cities, here("inst/extdata/eurovision_host_cities.csv"))
```

# Inspect Participants df

Claude assistance for wrangling code

## Contest Format

```{r}
# ── Contest Format Reference Data Frame ───────────────────────────────────────
# Built from historical knowledge of Eurovision format changes [1][2][6]

df_contest_format <- tibble(
  year = setdiff(1956:2026, 2020)
) |>
  mutate(
    format = case_when(

      # ── Grand Final only ────────────────────────────────────────────────────
      year %in% 1956:1992 ~ "Grand Final only",

      # ── Relegation / qualifier era ──────────────────────────────────────────
      # 1993: first relegation qualifier round introduced
      year == 1993        ~ "Relegation qualifier + Grand Final",

      # 1994–1995: formal relegation rule — bottom countries sat out
      year %in% 1994:1995 ~ "Relegation rule + Grand Final",

      # 1996: all countries participated in a qualifier round
      year == 1996        ~ "Open qualifier round + Grand Final",

      # 1997–2003: rolling points window determined automatic qualifiers
      year %in% 1997:2003 ~ "Rolling points relegation + Grand Final",

      # ── Semi-final era ───────────────────────────────────────────────────────
      # 2004–2007: one semi-final introduced
      year %in% 2004:2007 ~ "One semi-final + Grand Final",

      # 2008–present: two semi-finals
      year %in% 2008:2026 ~ "Two semi-finals + Grand Final",

      TRUE                ~ "Unknown"
    ),

    # ── Numeric flags for easy filtering ──────────────────────────────────────
    has_grand_final    = TRUE,
    has_semi_final_1   = year >= 2004,
    has_semi_final_2   = year >= 2008,
    has_relegation     = year %in% 1993:2003,

    # ── Number of live shows ──────────────────────────────────────────────────
    n_live_shows = case_when(
      year %in% 1956:2003 ~ 1L,
      year %in% 2004:2007 ~ 2L,
      year >= 2008        ~ 3L
    )
  )

# ── Save ──────────────────────────────────────────────────────────────────────
write_csv(df_contest_format, here("inst/extdata/eurovision_contest_format.csv"))
```

## Participants df

```{r}
# ── Post-processing: Split participants into two clean data frames ─────────────
raw <- participants

# ── Replace all NA-like strings with true NA ──────────────────────────────────
na_strings <- c("NA", "", "N/A", "–", "—")

clean_na <- function(df) {
  df |>
    mutate(across(
      everything(),
      ~ if_else(.x %in% na_strings, NA_character_, .x)
    ))
}

raw <- clean_na(raw)

# ═══════════════════════════════════════════════════════════════════════════════
# DATA FRAME 1 — Participant Info (year → conductor)
# ═══════════════════════════════════════════════════════════════════════════════

df_participants_info <- raw |>

  filter(str_detect(table_label, "Participants \\[1\\]")) |>

  select(year, country, broadcaster, artist, song, language,
         songwriter_s, conductor, language_s) |>

  mutate(language = coalesce(language, language_s)) |>
  select(-language_s) |>

  mutate(
    song = str_remove_all(song, '^"+|"+$'),
    song = str_squish(song)
  ) |>

  filter(!(is.na(artist) & is.na(song))) |>

  mutate(year = as.integer(year)) |>

  arrange(year, country)

# rewrite 
write_csv(df_participants_info, here("inst/extdata/eurovision_participants.csv"))

# ═══════════════════════════════════════════════════════════════════════════════
# DATA FRAME 2 — Semi-final Results (2004 onwards)
# Source: eurovision_data$results, filtered to semi-final sections only
# ═══════════════════════════════════════════════════════════════════════════════

# ── Apply same NA cleaning to results ─────────────────────────────────────────
results_clean <- results |>
  mutate(across(where(is.character), ~ if_else(.x %in% na_strings, NA_character_, .x)))

df_semifinal_results <- results_clean |>

  # ── Keep only semi-final years ─────────────────────────────────────────────
  filter(as.integer(year) >= 2004, as.integer(year) != 2020) |>

  # ── Keep only semi-final section tables ───────────────────────────────────
  # Exclude "Semi-final allocation draw" (the pot draw table)
  # Keep only labels matching "Semi-final" but NOT "allocation"
  filter(str_detect(table_label, regex("semi-final", ignore_case = TRUE))) |>
  filter(!str_detect(table_label, regex("allocation", ignore_case = TRUE))) |>

  # ── Keep only the FIRST table per semi-final section per year ─────────────
  # The first table (lowest index number in the label) is the results table.
  # Later tables with higher indices are the detailed voting breakdowns.
  # We extract the index number and keep only the minimum per year + section.
  mutate(
    # Extract the section name without the index: e.g. "Semi-final 1"
    section = str_extract(table_label, "Semi-final(?: \\d)?"),
    section = str_squish(section),
    # Extract the numeric index from the label e.g. [2] → 2
    tbl_index = as.integer(str_extract(table_label, "(?<=\\[)\\d+(?=\\])"))
  ) |>

  # Keep only the lowest-indexed table per year + section
  group_by(year, section) |>
  filter(tbl_index == min(tbl_index)) |>
  ungroup() |>

  # ── Drop the first NA row (Wikipedia table header artefact) ───────────────
  filter(!is.na(country)) |>

  # ── Select and clean core columns ─────────────────────────────────────────
  mutate(
    # Normalise section name: "Semi-final" → "Semi-final 1" for 2004–2007
    semi_final = if_else(
      section == "Semi-final",
      "Semi-final 1",
      section
    ),
    year = as.integer(year),

    # Clean song titles
    song  = str_remove_all(song, '^"+|"+$'),
    song  = str_squish(song),

    # Strip footnote markers from country
    country = str_remove_all(country, "[‡†\\*]"),
    country = str_squish(country),

    # r_o is running order (draw position); place and points are results
    running_order = suppressWarnings(as.integer(r_o)),
    place         = suppressWarnings(as.integer(place)),
    points        = suppressWarnings(as.numeric(points)),

    # Qualified: top 10 advance from each semi-final
    qualified = case_when(
      !is.na(place) & place <= 10L ~ TRUE,
      !is.na(place) & place >  10L ~ FALSE,
      TRUE                         ~ NA
    ),

    # Format label
    format = if_else(
      year %in% 2004:2007,
      "One semi-final + Grand Final",
      "Two semi-finals + Grand Final"
    )
  ) |>

  select(
    year,
    format,
    semi_final,
    running_order,
    country,
    artist,
    song,
    language,
    place,
    points,
    qualified
  ) |>

  arrange(year, semi_final, place)

# ── Save ───────────────────────────────────────────────────────────────────────
write_csv(df_semifinal_results, here("inst/extdata/eurovision_semifinal_results.csv"))
```

# Inspect Results df

use claude to help remove the semi results from the final results etc

also identify the number each country votes for another country ?

Can you please clean this data frame using R tidyverse-friendly language so that we only have the grand final results for contests from 1956 to 2026. Please create two clear data frames: one that contains the total votes per song/country for each year, and another that outlines the number of votes given by each country for each song/country per year (if that information is available) - the latter would likely be a list of many data frames - one for each year. Once you have that list of many data frames, please join into one large data frame, separating each contest by year. 

More importantly, separate by the table_label column. There are different "detailed voting results", "final", "semi-final", "semi-final allocation draw", "table" x. Please separate these accordingly and explain why these different tables exist. 

Ideally, I would like the following data frames: 

1. total votes for each semi final (eurovision_semifinal_results.csv) 
2. total votes for the grand final
3. votes per song for each semi final awarded by each country (a list of data frames per year)
4. votes per song for each grand final awarded by each country (a list of data frames per year) 
5. semi-final allocation draw data 

Outline any data that we have but do not have a home for in these data frames.

Notes: 
- First Eurovision (1956): Detailed voting data is incomplete. Only aggregate results may be available.
- Data Quality: EurovisionWorld is more reliable for historical data than Wikipedia's variable HTML structure.
- Semi-Finals: Semi-final data is typically only available from 2004 onwards when the contest expanded to multiple shows.
- Jury vs. Televote Split: From 2009 onwards, voting is split 50/50 between jury and public, making data more granular.

table_label                       n
results | Detailed voting results [1]	889	
results | Detailed voting results [2]	337	
results | Detailed voting results [3]	12	
results | Final [10]	127	
results | Final [11]	259	
results | Final [12]	190	
results | Final [13]	121	
results | Final [15]	187	
results | Final [16]	195	
results | Final [17]	195	
results | Final [18]	100	
results | Final [19]	91	
results | Final [1]	23	
results | Final [2]	96	
results | Final [4]	486	
results | Final [5]	110	
results | Final [6]	59	
results | Final [9]	133	
results | Semi-final 1 [2]	304	
results | Semi-final 1 [5]	322	
results | Semi-final 1 [6]	281	
results | Semi-final 1 [7]	195	
results | Semi-final 1 [8]	63	
results | Semi-final 1 [9]	62	
results | Semi-final 2 [10]	183	
results | Semi-final 2 [11]	136	
results | Semi-final 2 [12]	136	
results | Semi-final 2 [13]	53	
results | Semi-final 2 [14]	62	
results | Semi-final 2 [3]	311	
results | Semi-final 2 [7]	91	
results | Semi-final 2 [8]	156	
results | Semi-final 2 [9]	105	
results | Semi-final [1]	98	
results | Semi-final [3]	102	
results | Semi-final [4]	53	
results | Semi-final allocation draw [1]	18	
results | table 1	14	
results | table 2	14	
results | table 3	17	
results | table 4	5	

## 1956 [Table 1, 2, 3, 4]

We know: 
- 1956 did not disclose votes per country or total. 
- There were multiple songs per person but they only competed with one final song 

```{r}
results_1956 <- results %>% 
  # filter table 1 
  filter(table_label == "results | table 1") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # known winner of the contest 
  mutate(total_score = ifelse(artist == "Lys Assia" & song == '"Refrain"', 1, 0)) %>% 
  # remove extra songs per artist 
  mutate(
    order_s = case_when(
      song == '"De vogels van Holland"' ~ "1",
      song ==	'"Das alte Karussell"' ~ "2",
      song ==	'"Messieurs les noyés de la Seine"' ~ "3",
      song == '"Im Wartesaal zum großen Glück"' ~ "4",
      song == '"Le Temps perdu"' ~ "5",
      song ==	'"Ne crois pas"' ~ "6",
      song == '"Aprite le finestre"' ~ "7",
      song == '"Voorgoed voorbij"' ~ "8",
      song == '"Refrain"' ~ "9",
      song == '"Le Plus Beau Jour de ma vie"' ~ "10",
      song == '"So geht das jede Nacht"'~ "11",
      song == '"Il est là"' ~ "12",
      song == '"Les Amants de minuit"' ~ "13",
      song == '"Amami se vuoi"' ~ "14",
      .default = "check"
    ),
    order_s = as.numeric(order_s)
  ) 

# Table 2 is unnecessary as it only contains the order which we've just manually entered.
# Table 3 and 4 contains the commentary information that is already in the 'broadcasts` df. 
results_cleaned <- results %>% 
  # remove table 1, 2, 3, 4 
  filter(!str_detect(table_label, "table")) %>% 
  # rename awful col name 
  rename("total_score_1" = "supports_writing_mode_vertical_rl_mw_parser_output_ts_vertical_header_line_height_1_max_width_1em_padding_0_4em_vertical_align_bottom_width_1em_html_client_js_mw_parser_output_sortable_not_jquery_tablesorter_ts_vertical_header_not_unsortable_html_client_js_mw_parser_output_ts_vertical_header_header_sort_background_position_50_percent_4em_padding_right_0_4em_padding_top_21px_mw_parser_output_ts_vertical_header_is_valign_top_vertical_align_top_mw_parser_output_ts_vertical_header_is_valign_middle_vertical_align_middle_mw_parser_output_ts_vertical_header_is_normal_font_weight_normal_mw_parser_output_ts_vertical_header_display_inline_block_transform_rotate_180deg_writing_mode_vertical_rl_supports_writing_mode_sideways_lr_mw_parser_output_ts_vertical_header_transform_none_writing_mode_sideways_lr_total_score")
```

## 1957-1995 [Detailed Results 1]

```{r}
results_1957_to_1995 <- results_cleaned %>% 
  # filter Detailed voting results [1]
  filter(table_label == "results | Detailed voting results [1]") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>%
  # remove table_label
  select(-table_label) %>% 
# filter years without NAs for x1 and x2 
  filter(year %in% c(1957:1995))

results_1957_to_1995_detailed <- results_1957_to_1995 %>% 
  # remove totals and keep breakdown matricies 
  dplyr::select(-c(total_score_1, starts_with("voting"), x1)) %>% 
  # rename x2 to country 1 of matrix 
  rename("performing_country" = x2) %>% 
  # pivot countries 
  pivot_longer(
    cols = c(switzerland:ukraine), 
    names_to = "voting_country", 
    values_to = "score"
  )
# next step would be to add 0's if they were participants but received no votes, while NA if they did not participate and could not vote. 

# results_1957_to_1995_final <- 
  # results_1957_to_1995 %>% 
  # # remove detailed results 
  # dplyr::select(-c(x1, switzerland:macedonia, latvia:ukraine)) %>% 
  # # identify what's in the voting procedue columns
  # select(where(~!all(is.na(.))))

# 1997 to 2003 
results_1997_to_2003_1 <- results_cleaned %>% 
  # filter Detailed voting results [1]
  filter(table_label == "results | Detailed voting results [1]") %>% 
  select(year, total_score_1, starts_with("voting")) %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # filter so we can see where there ARE values in these columns
  filter(!(
    is.na(voting_procedure_used_100_percent_jury_vote_100_percent_televoting_1) &
    is.na(voting_procedure_used_100_percent_televoting_100_percent_jury_vote_1) & 
    is.na(voting_procedure_used_100_percent_televoting_100_percent_jury_vote_2)
  )) %>% 
  # they appear to be misaligned data 
  rename(
    x1 = voting_procedure_used_100_percent_jury_vote_100_percent_televoting_1,
    x2 = voting_procedure_used_100_percent_jury_vote_100_percent_televoting_2, 
    x3 = voting_procedure_used_100_percent_televoting_100_percent_jury_vote_1,
    x4 = voting_procedure_used_100_percent_televoting_100_percent_jury_vote_2
  ) %>% 
  # remove unneccessary columns
  select(-c(x1, x3)) %>% 
  # join x2 and x4 together
  mutate(performing_country = coalesce(x2, x4)) %>% 
  # remove old cols
  select(-c(x2, x4))

results_1997_to_2003_voting <- results_1997_to_2003_1 %>%
  # filter for voting procedure used data
  filter(str_detect(performing_country, "Voting"))

results_1997_to_2003_1 %>% 
  filter(!str_detect(performing_country, "Voting"))
```


```{r}
results_clean <- results %>%
  mutate(
    table_type = case_when(
      str_detect(table_label, "allocation draw") ~ "allocation_draw",
      str_detect(table_label, "Detailed voting results") ~ "detailed_voting",
      str_detect(table_label, "Final") ~ "final",
      str_detect(table_label, "Semi-final [12]") ~ "semifinal_summary",
      str_detect(table_label, "Semi-final \\d \\[") ~ "semifinal_detailed",
      str_detect(table_label, "^results \\| table") ~ "legacy_table",
      TRUE ~ "other"
    ),
    semi_final_number = case_when(
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  )

country_cols <- results_clean %>%
  select(where(is.numeric)) %>%
  select(-year, -r_o, -n, -no, -points, -number, 
         -starts_with("televote_"), -starts_with("jury_vote_"),
         -starts_with("combined_"), -starts_with("rest_"), 
         -semi_final_number) %>%
  colnames()

# ============================================================================
# EXTRACT VOTING DATA (LONG FORMAT) - FIXED
# ============================================================================

voting_data <- results_clean %>%
  select(
    year, table_label, table_type, country, 
    song, semi_final_number, artist, 
    all_of(country_cols)
  ) %>%
  pivot_longer(
    cols = all_of(country_cols),
    names_to = "country_voting",
    values_to = "votes_awarded"
  ) %>%
  filter(!is.na(votes_awarded), votes_awarded > 0, !is.na(country)) %>%
  mutate(
    country_voting = str_to_title(str_replace_all(country_voting, "_", " ")),
    country_performing = country
  ) %>%
  select(year, table_label, table_type, semi_final_number, 
         country_performing, song, country_voting, votes_awarded)

```

```{r}

head(voting_data, 20)
cat("\n✓ Voting data created successfully!\n")
cat("Rows:", nrow(voting_data), "\n")

# ============================================================================
# OUTPUT 1: SEMIFINAL TOTALS
# ============================================================================

eurovision_semifinal_totals <- voting_data %>%
  filter(table_type %in% c("semifinal_summary", "other")) %>%
  mutate(
    semi_final_number = case_when(
      !is.na(semi_final_number) ~ semi_final_number,
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(semi_final_number)) %>%
  group_by(year, semi_final_number, song, country_performing) %>%
  summarise(
    total_votes = sum(votes_awarded, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(country = country_performing) %>%
  select(year, semi_final_number, song, country, total_votes) %>%
  arrange(year, semi_final_number, desc(total_votes))

head(eurovision_semifinal_totals, 20)
cat("\nSemi-final totals rows:", nrow(eurovision_semifinal_totals), "\n")

# ============================================================================
# OUTPUT 2: FINAL TOTALS
# ============================================================================

eurovision_final_totals <- voting_data %>%
  filter(table_type == "final") %>%
  group_by(year, song, country_performing) %>%
  summarise(
    total_votes = sum(votes_awarded, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(country = country_performing) %>%
  select(year, song, country, total_votes) %>%
  arrange(year, desc(total_votes))

head(eurovision_final_totals, 20)
cat("\nFinal totals rows:", nrow(eurovision_final_totals), "\n")

# ============================================================================
# OUTPUT 3: SEMIFINAL VOTING BY COUNTRY (LIST)
# ============================================================================

eurovision_semifinal_voting_by_country <- voting_data %>%
  filter(table_type %in% c("semifinal_summary", "other")) %>%
  mutate(
    semi_final_number = case_when(
      !is.na(semi_final_number) ~ semi_final_number,
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(semi_final_number)) %>%
  select(year, semi_final_number, song, country_performing, 
         country_voting, votes_awarded) %>%
  arrange(year, semi_final_number) %>%
  nest(data = -year) %>%
  deframe()

cat("\nSemi-final voting by country: list with", 
    length(eurovision_semifinal_voting_by_country), "years\n")
if (length(eurovision_semifinal_voting_by_country) > 0) {
  head(eurovision_semifinal_voting_by_country[[1]], 15)
}

# ============================================================================
# OUTPUT 4: FINAL VOTING BY COUNTRY (LIST)
# ============================================================================

eurovision_final_voting_by_country <- voting_data %>%
  filter(table_type == "final") %>%
  select(year, song, country_performing, country_voting, votes_awarded) %>%
  arrange(year) %>%
  nest(data = -year) %>%
  deframe()

cat("\nFinal voting by country: list with", 
    length(eurovision_final_voting_by_country), "years\n")
if (length(eurovision_final_voting_by_country) > 0) {
  head(eurovision_final_voting_by_country[[1]], 15)
}

# ============================================================================
# OUTPUT 5: SEMIFINAL ALLOCATION DRAW
# ============================================================================

eurovision_semifinal_allocation_draw <- results_clean %>%
  filter(table_type == "allocation_draw") %>%
  select(year, country, semi_final_assigned = semi_final_number, 
         broadcaster, language) %>%
  filter(!is.na(semi_final_assigned), !is.na(country)) %>%
  arrange(year, semi_final_assigned, country)

head(eurovision_semifinal_allocation_draw, 20)
cat("\nSemi-final allocation draw rows:", nrow(eurovision_semifinal_allocation_draw), "\n")

# ============================================================================
# DATA QUALITY CHECKS
# ============================================================================

cat("\n\n=== DATA QUALITY REPORT ===\n\n")

cat("📈 SUMMARY STATISTICS:\n")
cat("✓ Semi-final totals:", nrow(eurovision_semifinal_totals), "rows\n")
cat("✓ Final totals:", nrow(eurovision_final_totals), "rows\n")
cat("✓ Semi-final voting years:", length(eurovision_semifinal_voting_by_country), "\n")
cat("✓ Final voting years:", length(eurovision_final_voting_by_country), "\n")
cat("✓ Allocation draw entries:", nrow(eurovision_semifinal_allocation_draw), "\n")

```

# Inspect new df

```{r}
df_contest_format     <- read_csv(here("inst/extdata/eurovision_contest_format.csv"))
df_participants_info  <- read_csv(here("inst/extdata/eurovision_participants.csv"))
df_semifinal_results  <- read_csv(here("inst/extdata/eurovision_semifinal_results.csv"))
```

