---
title: "Data Wrangling"
format: html
---

To Do: 
[] Add place and total score to participants df 
[] Make results df in long format 
[] Make all df country names lower case
[] Create function to make country names sentence case with and as lower case for visualisation

```{r}
library(tidyverse)
library(here)
library(janitor)

host_cities   <- read_csv(here("inst/extdata/eurovision_host_cities.csv"))
participants  <- read_csv(
  here("inst/extdata/eurovision_participants.csv"),
  show_col_types = FALSE,
  col_types       = cols(.default = col_character())
)
results       <- read_csv(here("inst/extdata/eurovision_results.csv"))
broadcasts    <- read_csv(here("inst/extdata/eurovision_broadcasts.csv"))
```

# Inspect Host Cities df

Manual inspection and edits

```{r}
host_cities <- host_cities %>% 
  mutate(
    venue = case_when(
      year == 1965 ~ "Sala di Concerto della RAI",
      .default = venue
    ),
    host_city = case_when(
      year == 1965 ~ "Naples, Italy",
      year == 1974 ~ "Brighton, United Kingdom",
      year == 2010 ~ "Oslo, Norway",
      year == 2013 ~ "Malmö, Sweden",
      .default = host_city
    )
  )

# rewrite over 
write_csv(host_cities, here("inst/extdata/eurovision_host_cities.csv"))
```

# Inspect Participants df

Claude assistance for wrangling code

## Contest Format

```{r}
# ── Contest Format Reference Data Frame ───────────────────────────────────────
# Built from historical knowledge of Eurovision format changes [1][2][6]

df_contest_format <- tibble(
  year = setdiff(1956:2026, 2020)
) |>
  mutate(
    format = case_when(

      # ── Grand Final only ────────────────────────────────────────────────────
      year %in% 1956:1992 ~ "Grand Final only",

      # ── Relegation / qualifier era ──────────────────────────────────────────
      # 1993: first relegation qualifier round introduced
      year == 1993        ~ "Relegation qualifier + Grand Final",

      # 1994–1995: formal relegation rule — bottom countries sat out
      year %in% 1994:1995 ~ "Relegation rule + Grand Final",

      # 1996: all countries participated in a qualifier round
      year == 1996        ~ "Open qualifier round + Grand Final",

      # 1997–2003: rolling points window determined automatic qualifiers
      year %in% 1997:2003 ~ "Rolling points relegation + Grand Final",

      # ── Semi-final era ───────────────────────────────────────────────────────
      # 2004–2007: one semi-final introduced
      year %in% 2004:2007 ~ "One semi-final + Grand Final",

      # 2008–present: two semi-finals
      year %in% 2008:2026 ~ "Two semi-finals + Grand Final",

      TRUE                ~ "Unknown"
    ),

    # ── Numeric flags for easy filtering ──────────────────────────────────────
    has_grand_final    = TRUE,
    has_semi_final_1   = year >= 2004,
    has_semi_final_2   = year >= 2008,
    has_relegation     = year %in% 1993:2003,

    # ── Number of live shows ──────────────────────────────────────────────────
    n_live_shows = case_when(
      year %in% 1956:2003 ~ 1L,
      year %in% 2004:2007 ~ 2L,
      year >= 2008        ~ 3L
    )
  )

# ── Save ──────────────────────────────────────────────────────────────────────
write_csv(df_contest_format, here("inst/extdata/eurovision_contest_format.csv"))
```

## Participants df

```{r}
# ── Post-processing: Split participants into two clean data frames ─────────────
raw <- participants

# ── Replace all NA-like strings with true NA ──────────────────────────────────
na_strings <- c("NA", "", "N/A", "–", "—")

clean_na <- function(df) {
  df |>
    mutate(across(
      everything(),
      ~ if_else(.x %in% na_strings, NA_character_, .x)
    ))
}

raw <- clean_na(raw)

# ═══════════════════════════════════════════════════════════════════════════════
# DATA FRAME 1 — Participant Info (year → conductor)
# ═══════════════════════════════════════════════════════════════════════════════

df_participants_info <- raw |>

  filter(str_detect(table_label, "Participants \\[1\\]")) |>

  select(year, country, broadcaster, artist, song, language,
         songwriter_s, conductor, language_s) |>

  mutate(language = coalesce(language, language_s)) |>
  select(-language_s) |>

  mutate(
    song = str_remove_all(song, '^"+|"+$'),
    song = str_squish(song)
  ) |>

  filter(!(is.na(artist) & is.na(song))) |>

  mutate(year = as.integer(year)) |>

  arrange(year, country)

# rewrite 
write_csv(df_participants_info, here("inst/extdata/eurovision_participants.csv"))

# ═══════════════════════════════════════════════════════════════════════════════
# DATA FRAME 2 — Semi-final Results (2004 onwards)
# Source: eurovision_data$results, filtered to semi-final sections only
# ═══════════════════════════════════════════════════════════════════════════════

# ── Apply same NA cleaning to results ─────────────────────────────────────────
results_clean <- results |>
  mutate(across(where(is.character), ~ if_else(.x %in% na_strings, NA_character_, .x)))

df_semifinal_results <- results_clean |>

  # ── Keep only semi-final years ─────────────────────────────────────────────
  filter(as.integer(year) >= 2004, as.integer(year) != 2020) |>

  # ── Keep only semi-final section tables ───────────────────────────────────
  # Exclude "Semi-final allocation draw" (the pot draw table)
  # Keep only labels matching "Semi-final" but NOT "allocation"
  filter(str_detect(table_label, regex("semi-final", ignore_case = TRUE))) |>
  filter(!str_detect(table_label, regex("allocation", ignore_case = TRUE))) |>

  # ── Keep only the FIRST table per semi-final section per year ─────────────
  # The first table (lowest index number in the label) is the results table.
  # Later tables with higher indices are the detailed voting breakdowns.
  # We extract the index number and keep only the minimum per year + section.
  mutate(
    # Extract the section name without the index: e.g. "Semi-final 1"
    section = str_extract(table_label, "Semi-final(?: \\d)?"),
    section = str_squish(section),
    # Extract the numeric index from the label e.g. [2] → 2
    tbl_index = as.integer(str_extract(table_label, "(?<=\\[)\\d+(?=\\])"))
  ) |>

  # Keep only the lowest-indexed table per year + section
  group_by(year, section) |>
  filter(tbl_index == min(tbl_index)) |>
  ungroup() |>

  # ── Drop the first NA row (Wikipedia table header artefact) ───────────────
  filter(!is.na(country)) |>

  # ── Select and clean core columns ─────────────────────────────────────────
  mutate(
    # Normalise section name: "Semi-final" → "Semi-final 1" for 2004–2007
    semi_final = if_else(
      section == "Semi-final",
      "Semi-final 1",
      section
    ),
    year = as.integer(year),

    # Clean song titles
    song  = str_remove_all(song, '^"+|"+$'),
    song  = str_squish(song),

    # Strip footnote markers from country
    country = str_remove_all(country, "[‡†\\*]"),
    country = str_squish(country),

    # r_o is running order (draw position); place and points are results
    running_order = suppressWarnings(as.integer(r_o)),
    place         = suppressWarnings(as.integer(place)),
    points        = suppressWarnings(as.numeric(points)),

    # Qualified: top 10 advance from each semi-final
    qualified = case_when(
      !is.na(place) & place <= 10L ~ TRUE,
      !is.na(place) & place >  10L ~ FALSE,
      TRUE                         ~ NA
    ),

    # Format label
    format = if_else(
      year %in% 2004:2007,
      "One semi-final + Grand Final",
      "Two semi-finals + Grand Final"
    )
  ) |>

  select(
    year,
    format,
    semi_final,
    running_order,
    country,
    artist,
    song,
    language,
    place,
    points,
    qualified
  ) |>

  arrange(year, semi_final, place)

# ── Save ───────────────────────────────────────────────────────────────────────
write_csv(df_semifinal_results, here("inst/extdata/eurovision_semifinal_results.csv"))
```

# Inspect Results df

use claude to help remove the semi results from the final results etc

also identify the number each country votes for another country ?

Can you please clean this data frame using R tidyverse-friendly language so that we only have the grand final results for contests from 1956 to 2026. Please create two clear data frames: one that contains the total votes per song/country for each year, and another that outlines the number of votes given by each country for each song/country per year (if that information is available) - the latter would likely be a list of many data frames - one for each year. Once you have that list of many data frames, please join into one large data frame, separating each contest by year. 

More importantly, separate by the table_label column. There are different "detailed voting results", "final", "semi-final", "semi-final allocation draw", "table" x. Please separate these accordingly and explain why these different tables exist. 

Ideally, I would like the following data frames: 

1. total votes for each semi final (eurovision_semifinal_results.csv) 
2. total votes for the grand final
3. votes per song for each semi final awarded by each country (a list of data frames per year)
4. votes per song for each grand final awarded by each country (a list of data frames per year) 
5. semi-final allocation draw data 

Outline any data that we have but do not have a home for in these data frames.

Notes: 
- First Eurovision (1956): Detailed voting data is incomplete. Only aggregate results may be available.
- Data Quality: EurovisionWorld is more reliable for historical data than Wikipedia's variable HTML structure.
- Semi-Finals: Semi-final data is typically only available from 2004 onwards when the contest expanded to multiple shows.
- Jury vs. Televote Split: From 2009 onwards, voting is split 50/50 between jury and public, making data more granular.

table_label                       n
results | Detailed voting results [1]	889	
results | Detailed voting results [2]	337	
results | Detailed voting results [3]	12	
results | Final [10]	127	
results | Final [11]	259	
results | Final [12]	190	
results | Final [13]	121	
results | Final [15]	187	
results | Final [16]	195	
results | Final [17]	195	
results | Final [18]	100	
results | Final [19]	91	
results | Final [1]	23	
results | Final [2]	96	
results | Final [4]	486	
results | Final [5]	110	
results | Final [6]	59	
results | Final [9]	133	
results | Semi-final 1 [2]	304	
results | Semi-final 1 [5]	322	
results | Semi-final 1 [6]	281	
results | Semi-final 1 [7]	195	
results | Semi-final 1 [8]	63	
results | Semi-final 1 [9]	62	
results | Semi-final 2 [10]	183	
results | Semi-final 2 [11]	136	
results | Semi-final 2 [12]	136	
results | Semi-final 2 [13]	53	
results | Semi-final 2 [14]	62	
results | Semi-final 2 [3]	311	
results | Semi-final 2 [7]	91	
results | Semi-final 2 [8]	156	
results | Semi-final 2 [9]	105	
results | Semi-final [1]	98	
results | Semi-final [3]	102	
results | Semi-final [4]	53	
results | Semi-final allocation draw [1]	18	
results | table 1	14	
results | table 2	14	
results | table 3	17	
results | table 4	5	

## 1956 [Table 1, 2, 3, 4]

We know: 
- 1956 did not disclose votes per country or total. 
- There were multiple songs per person but they only competed with one final song 

```{r}
results_1956 <- results %>% 
  # filter table 1 
  filter(table_label == "results | table 1") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # known winner of the contest 
  mutate(total_score = ifelse(artist == "Lys Assia" & song == '"Refrain"', 1, 0)) %>% 
  # remove extra songs per artist 
  mutate(
    order_s = case_when(
      song == '"De vogels van Holland"' ~ "1",
      song ==	'"Das alte Karussell"' ~ "2",
      song ==	'"Messieurs les noyés de la Seine"' ~ "3",
      song == '"Im Wartesaal zum großen Glück"' ~ "4",
      song == '"Le Temps perdu"' ~ "5",
      song ==	'"Ne crois pas"' ~ "6",
      song == '"Aprite le finestre"' ~ "7",
      song == '"Voorgoed voorbij"' ~ "8",
      song == '"Refrain"' ~ "9",
      song == '"Le Plus Beau Jour de ma vie"' ~ "10",
      song == '"So geht das jede Nacht"'~ "11",
      song == '"Il est là"' ~ "12",
      song == '"Les Amants de minuit"' ~ "13",
      song == '"Amami se vuoi"' ~ "14",
      .default = "check"
    ),
    order_s = as.numeric(order_s)
  )  %>% 
  # lower case performing_country var too 
  mutate(country = str_to_lower(str_replace_all(country, " ", "_"))) %>% 
  # add col re final or semi
  mutate(str = "final")

write_csv(results_1956, here("inst/extdata/eurovision_1956.csv"))
```

## Results -> Clean

```{r}
# Table 2 is unnecessary as it only contains the order which we've just manually entered.
# Table 3 and 4 contains the commentary information that is already in the 'broadcasts` df. 
results_cleaned <- results %>% 
  # remove table 1, 2, 3, 4 
  filter(!str_detect(table_label, "table")) %>% 
  # rename awful col name 
  rename("total_score_1" = "supports_writing_mode_vertical_rl_mw_parser_output_ts_vertical_header_line_height_1_max_width_1em_padding_0_4em_vertical_align_bottom_width_1em_html_client_js_mw_parser_output_sortable_not_jquery_tablesorter_ts_vertical_header_not_unsortable_html_client_js_mw_parser_output_ts_vertical_header_header_sort_background_position_50_percent_4em_padding_right_0_4em_padding_top_21px_mw_parser_output_ts_vertical_header_is_valign_top_vertical_align_top_mw_parser_output_ts_vertical_header_is_valign_middle_vertical_align_middle_mw_parser_output_ts_vertical_header_is_normal_font_weight_normal_mw_parser_output_ts_vertical_header_display_inline_block_transform_rotate_180deg_writing_mode_vertical_rl_supports_writing_mode_sideways_lr_mw_parser_output_ts_vertical_header_transform_none_writing_mode_sideways_lr_total_score") %>% 
  # lower case performing_country var too 
  mutate(
    country = str_to_lower(str_replace_all(country, " ", "_")),
    x2 = str_to_lower(str_replace_all(x2, " ", "_"))
  )
```

## 1957-1995 [Detailed Results 1]

### Function: clean_matrix

```{r}
clean_matrix <- function(df){

  df %>% 
    # remove columns that are only NA 
    select(where(~ !all(is.na(.x)))) %>% 
    # pivot countries 
    pivot_longer(
      cols = -c(year, performing_country), 
      names_to = "voting_country", 
      values_to = "score"
    ) %>% 
    # # group by year 
    # group_by(year) %>%
    # # keep only rows where the country appears as both a performer and a voter that year
    # filter(performing_country %in% voting_country, voting_country %in% performing_country) %>% 
    # ungroup
    # ungroup() %>% 
    # replace NA with 0 
    replace_na(replace = list(score = 0))
  
}
```

### Data processing 

```{r}
results_1957_to_1995_filter <- results_cleaned %>% 
  # filter Detailed voting results [1]
  filter(table_label == "results | Detailed voting results [1]") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>%
  # remove table_label
  select(-table_label) %>% 
# filter years without NAs for x1 and x2 
  filter(year %in% c(1957:1995))

results_1957_to_1995_detailed <-
  results_1957_to_1995_filter %>% 
  # remove totals and keep breakdown matricies 
  dplyr::select(-c(total_score_1, starts_with("voting"), x1)) %>% 
  # rename x2 to country 1 of matrix 
  rename("performing_country" = x2) %>%
  # apply clean matrix
  clean_matrix() %>% 
  # remove anomaly in 1982 vote structure matrix
  filter(!c(year==1982 & is.na(performing_country)))

results_1957_to_1995_total <- results_1957_to_1995_filter %>%
  # rename x2 to country 1 of matrix 
  rename("performing_country" = x2) %>%
  # remove detailed results
  select(-c(x1, switzerland:macedonia, latvia:ukraine)) %>%
  # identify what's in the voting procedue columns
  select(where(~!all(is.na(.)))) %>% 
  # rename score
  rename(total_score = total_score_1)

# join back to one table
results_1957_to_1995 <- results_1957_to_1995_detailed %>% 
  # pivot wider again 
  pivot_wider(
    id_cols = c(year, performing_country),
    names_from = voting_country,
    values_from = score
  ) %>% 
  # join total score
  left_join(results_1957_to_1995_total, by = c("year", "performing_country")) %>% 
  # add col re final or semi
  mutate(str = "final")
```

### Add place 

```{r}
results_1957_to_1995 <- results_1957_to_1995 %>% 
  # group by year 
  group_by(year) %>% 
  # calculate place
  mutate(place = min_rank(desc(as.numeric(total_score)))) %>% 
  # ungroup 
  ungroup() %>% 
  # reorder
  select(year, performing_country, total_score, place, everything()) 

write_csv(results_1957_to_1995, here("inst/extdata/eurovision_1957_to_1995.csv"))
```

## 1996

Tables: 

1. results | Final [1] - final results with total points and place 
2. Detailed voting results [2] - detailed matrix of qualifying round votes
3. Detailed voting results [3] - distribution of 12 points awarded at qualifying round 
4. results | Final [4] - detailed matrix of final votes 
5. results | Final [5] - distribution of 12 points awarded at final

The distribution of 12 points is unnecessary as this can be obtained from the matrix

```{r}
results_1996_filter <- results_cleaned %>% 
  # filter to 1996
  filter(year == "1996") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.))))
```

### Qualifying

```{r}
results_1996_detailed_qual <- results_1996_filter %>% 
  # filter for qualifying matrix 
  filter(table_label == "results | Detailed voting results [2]") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # remove totals and keep breakdown matricies 
  dplyr::select(-c(total_score_1, table_label, x1)) %>% 
  # rename x2 to country 1 of matrix 
  rename("performing_country" = x2) %>%
  # apply clean matrix
  clean_matrix()

results_1996_total_qual <- results_1996_filter %>%
  # filter for qualifying matrix 
  filter(table_label == "results | Detailed voting results [2]") %>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # keep only relevant cols
  select(year, performing_country = x2, total_score = total_score_1) 

# join back to one table
results_1996_qual <- results_1996_detailed_qual %>% 
  # pivot wider again 
  pivot_wider(
    id_cols = c(year, performing_country),
    names_from = voting_country,
    values_from = score
  ) %>% 
  # join total score
  left_join(results_1996_total_qual, by = c("year", "performing_country")) %>% 
  # reorder
  select(year, performing_country, total_score, everything()) %>% 
  # add col re final or semi
  mutate(str = "qual")
```

### Final

```{r}
results_1996_points <- results_1996_filter %>% 
  # filter to appropriate table 
  filter(table_label == "results | Final [1]")%>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # remove unnecessary cols 
  select(year, performing_country = country, total_score = points, place)

results_1996_detailed_final <- results_1996_filter %>% 
  # filter to appropriate table 
  filter(table_label == "results | Final [4]")%>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # remove totals and keep breakdown matricies 
  dplyr::select(-c(table_label, total_score, x1)) %>% 
  # rename x2 to country 1 of matrix 
  rename("performing_country" = x2) %>%
  # apply clean matrix
  clean_matrix()

results_1996_total_final <- results_1996_filter %>%
  # filter for qualifying matrix 
  filter(table_label == "results | Final [4]")%>% 
  # remove NA columns 
  select(where(~!all(is.na(.)))) %>% 
  # keep only relevant cols
  select(year, performing_country = x2, total_score) 

# join back to one table
results_1996_final <- results_1996_detailed_final %>% 
  # pivot wider again 
  pivot_wider(
    id_cols = c(year, performing_country),
    names_from = voting_country,
    values_from = score
  ) %>% 
  # join total score
  left_join(results_1996_total_final, by = c("year", "performing_country")) %>% 
  # reorder
  select(year, performing_country, total_score, everything()) %>% 
  # add col re final or semi
  mutate(str = "final", total_score = as.numeric(total_score))

results_1996 <- results_1996_points %>% 
  # join
  left_join(results_1996_final, by = c("year", "performing_country", "total_score"))

write_csv(results_1996, here("inst/extdata/eurovision_1996.csv"))
```

### 12 points

Distribution of 12 points 

```{r}
# results_1996_filter %>% 
#   filter(table_label == "results | Detailed voting results [3]")%>% 
#   # remove NA columns 
#   select(where(~!all(is.na(.))))
# results_1996_filter %>% 
#   filter(table_label == "results | Final [5]")%>% 
#   # remove NA columns 
#   select(where(~!all(is.na(.))))
```

## 1997-2003

Two key data frames: 

1. Detailed voting results [1]: voting matrix
2. Detailed voting results [2]: distribution of 12 points awarded

Important note about 1997:

For the first time however, as part of a trial held by the contest organisers, televoting was used to determine the points from five of the participating countries.[5][38] In these countries viewers had a total of five minutes to register their vote by calling one of twenty-four different telephone numbers to represent the twenty-five competing entries except that which represented their own country. Once the voting phone lines were opened following the performance of the last competing entry, a video recap containing short clips of each competing entry with the accompanying phone number for voting was shown in order to aid viewers during the voting window.[40] In those countries which opted to use televoting to determine their points a jury was still required which would function as a back-up in case technical failure prevented the televote results from being used. The composition of the back-up juries in these countries was identical to the juries in the other countries with regards to profession, gender and age.[41]

100% televoting: Austria, Switzerland, Germany, Sweden, United Kingdom
100% jury voting: Cyprus, Turkey, Norway, Ireland, Slovenia, Netherlands, Italy, Spain, Poland, Estonia, Bosnia and Herzegovina, Portugal, Greece, Malta, Hungary, Russia, Denmark, France, Croatia, Iceland 

```{r}
results_1997_to_2003_filter <- results_cleaned %>% 
  # filter for years between 1997 to 2003
  filter(year %in% c(1997:2003))

results_1997_to_2003_filter %>% 
  summarise(n = n(), .by = "table_label")
```

### Data processing

```{r}
results_1997_to_2003_filter_1 <- results_1997_to_2003_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Detailed voting results [1]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.)))) %>% 
  # remove unnecessary cols
  select(
    -c(
      table_label, x1,
      voting_procedure_used_100_percent_jury_vote_100_percent_televoting_1, 
      voting_procedure_used_100_percent_televoting_100_percent_jury_vote_1
    )
  ) %>% 
  # filter where total score is NA (it's the column that outlines voting procedure used 
  # but when scraped from online it didn't capture the coloured boxes)
  filter(!is.na(total_score_1)) %>% 
  # rename percent voting cols
  rename(
    voting_jury_televote_1 = voting_procedure_used_100_percent_jury_vote_100_percent_televoting_2,
    voting_jury_televote_2 = voting_procedure_used_100_percent_televoting_100_percent_jury_vote_2,
    total_score = total_score_1
  ) %>% 
  # join all performing_country cols together 
  unite("performing_country", c(voting_jury_televote_1, voting_jury_televote_2, x2)) %>% 
  # remove na characters
  mutate(
    performing_country = gsub("_NA", "", performing_country),
    performing_country = gsub("NA_", "", performing_country),
    performing_country = str_to_lower(str_replace_all(performing_country, " ", "_")),
    total_score = as.numeric(total_score),
    str = "final"
  ) 

results_1997_to_2003_detailed <- 
  results_1997_to_2003_filter_1 %>%
  # remove cols
  select(-c(total_score, str)) %>% 
  # apply clean matrix
  clean_matrix() 

results_1997_to_2003 <- results_1997_to_2003_filter_1 %>% 
  # select cols
  select(year, performing_country, total_score, str) %>% 
  # join to clean detailed long format from matrix
  left_join(
    results_1997_to_2003_detailed %>% 
            # pivot wider again 
      pivot_wider(
        id_cols = c(year, performing_country),
        names_from = voting_country,
        values_from = score
      ) , by = c("year", "performing_country")
    ) %>% 
  # ungroup
  ungroup() %>% 
  # group by year 
  group_by(year) %>% 
  # calculate place
  mutate(place = min_rank(desc(as.numeric(total_score)))) %>% 
  # ungroup 
  ungroup() %>% 
  # reorder
  select(year, performing_country, total_score, place, everything()) 

write_csv(results_1997_to_2003, here("inst/extdata/eurovision_1997_to_2003.csv"))
```

### Add voting type 

Manual entry of which countries were jury and which were televote

Separate df for now

```{r}
# results_1997_to_2003 %>% 
#   mutate(
#     voting_str = case_when(
#       # before doing this, restructure so that it remains a long df 
#       # change format for all above as well 
#       # add cols to above years so that voting_str == "jury_100"
#       # then manually go through each of these years and state jury_100 or televote_100
#       # years beyond will be a mix of proportions with jury_televote e.g. jury_50_televote_50
#     )
#   )
```


## 2004-X

1. Semi-final [1]: 
2. Final [2]: 
3. Semi-final [3]: 
4. Semi-final [4]: 
5. Final [5]: 
6. Final [6]: 

```{r}
results_2004_to_2007_filter <- results_cleaned %>% 
  # filter for years between 2004 to 2007
  filter(year %in% c(2004:2007)) 

results_2004_to_2007_filter %>% 
  summarise(n = n(), .by = c("table_label", "year"))
```

### Tables 

```{r}
# table 1: semi final songs - unnecessary
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Semi-final [1]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 2: songs in final - unnecessary
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Final [2]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 3: semi-final scoring 
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Semi-final [3]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 4: semi-final scoring 12 points - unnecessary
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Semi-final [4]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 5: final scoring
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Final [5]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 6: final scoring 12 points - unnecessary
results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Final [6]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))
```

### Qualifying

```{r}
results_2004_to_2007_filter_qual_1 <- results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Semi-final [3]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.)))) %>% 
  # remove unnecessary cols
  select(
    -c(
      table_label,
      voting_procedure_used_100_percent_televoting_100_percent_jury_vote_1
    )
  ) %>% 
  # filter where total score is NA (it's the column that outlines voting procedure used 
  # but when scraped from online it didn't capture the coloured boxes)
  filter(!is.na(total_score_1)) %>% 
  # rename percent voting cols
  rename(
    performing_country = voting_procedure_used_100_percent_televoting_100_percent_jury_vote_2,
    total_score = total_score_1
  ) %>% 
  # adjust cols
  mutate(
    performing_country = str_to_lower(str_replace_all(performing_country, " ", "_")),
    total_score = as.numeric(total_score),
    str = "qual"
  ) 

results_2004_to_2007_detailed_qual <- 
  results_2004_to_2007_filter_qual_1 %>%
  # remove cols
  select(-c(total_score, str)) %>% 
  # apply clean matrix
  clean_matrix() 

results_2004_to_2007_qual <- results_2004_to_2007_filter_qual_1 %>% 
  # select cols
  select(year, performing_country, total_score, str) %>% 
  # join to clean detailed long format from matrix
  left_join(
    results_2004_to_2007_detailed_qual %>% 
            # pivot wider again 
      pivot_wider(
        id_cols = c(year, performing_country),
        names_from = voting_country,
        values_from = score
      ) , by = c("year", "performing_country")
    ) %>% 
  # ungroup
  ungroup() %>% 
  # group by year 
  group_by(year) %>% 
  # calculate place
  mutate(place = min_rank(desc(as.numeric(total_score)))) %>% 
  # ungroup 
  ungroup() %>% 
  # reorder
  select(year, performing_country, total_score, place, everything()) 
```

### Final

```{r}
results_2004_to_2007_filter_final_1 <- results_2004_to_2007_filter %>% 
  # filter to appropriate table
  filter(table_label == "results | Final [5]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.)))) %>% 
  # remove unnecessary cols
  select(
    -c(
      table_label,
      voting_procedure_used_100_percent_televoting_100_percent_jury_vote_1,
      voting_procedure_used_100_percent_televoting_1
    )
  ) %>% 
  # filter where total score is NA (it's the column that outlines voting procedure used 
  # but when scraped from online it didn't capture the coloured boxes)
  filter(!is.na(total_score)) %>% 
  # join all performing_country cols together 
  unite(
    "performing_country", 
    c(
      voting_procedure_used_100_percent_televoting_100_percent_jury_vote_2, 
      voting_procedure_used_100_percent_televoting_2
    ) 
  ) %>% 
  # remove na characters
  mutate(
    performing_country = gsub("_NA", "", performing_country),
    performing_country = gsub("NA_", "", performing_country),
    performing_country = str_to_lower(str_replace_all(performing_country, " ", "_")),
    total_score = as.numeric(total_score),
    str = "final"
  ) 

results_2004_to_2007_detailed_final <- 
  results_2004_to_2007_filter_final_1 %>%
  # remove cols
  select(-c(total_score, str)) %>% 
  # apply clean matrix
  clean_matrix() 

results_2004_to_2007_final <- results_2004_to_2007_filter_final_1 %>% 
  # select cols
  select(year, performing_country, total_score, str) %>% 
  # join to clean detailed long format from matrix
  left_join(
    results_2004_to_2007_detailed_final %>% 
            # pivot wider again 
      pivot_wider(
        id_cols = c(year, performing_country),
        names_from = voting_country,
        values_from = score
      ) , by = c("year", "performing_country")
    ) %>% 
  # ungroup
  ungroup() %>% 
  # group by year 
  group_by(year) %>% 
  # calculate place
  mutate(place = min_rank(desc(as.numeric(total_score)))) %>% 
  # ungroup 
  ungroup() %>% 
  # reorder
  select(year, performing_country, total_score, place, everything()) 
```

## 2008-2026

Known: 

- qualification now termed as semi-final 1 and semi-final 2
- 2020 eurovision did not occur 
- 2024 Joost Klein was disqualified from grand final 

```{r}
results_2008_to_2026_filter <- results_cleaned %>% 
  # filter for years between 1997 to 2003
  filter(year %in% c(2008:2026)) 

results_2008_to_2026_filter %>% 
  summarise(n = n(), .by = "table_label")

results_2008_to_2026_filter %>% 
  summarise(years = paste0(sort(unique(year)), collapse = ", "), .by = "table_label") 
```

results | Semi-final allocation draw [1]: every year 
results | Semi-final 1 [2]:  every year 
results | Semi-final 2 [3]:  every year 
results | Final [4]:  every year 
results | Semi-final 1 [5]:  every year 
results | Semi-final 1 [6]:  every year 
-
results | Semi-final 2 [7]: 
results | Semi-final 2 [8]: 
results | Final [9]: 
results | Final [10]: 
results | Final [11]: 
-
results | Semi-final 1 [7]: 
results | Semi-final 2 [9]: 
results | Semi-final 2 [10]: 
results | Final [12]: 
results | Final [13]: 
-
results | Semi-final 1 [8]: 
results | Semi-final 1 [9]: 
results | Semi-final 2 [11]: 
results | Semi-final 2 [12]: 
results | Semi-final 2 [13]: 
results | Semi-final 2 [14]: 
results | Final [15]: 
results | Final [16]: 
results | Final [17]: 
results | Final [18]: 
results | Final [19]: 

### Tables

```{r}
# table 1: allocation of semi-finals (2008-2026) - unnecessary 
results_2008_to_2026_filter %>% 
  # filter for table 
  filter(table_label=="results | Semi-final allocation draw [1]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 2: semi-final 1: songs, artist, overall points and placing - unnecessary 
results_2008_to_2026_filter %>% 
  # filter for table 
  filter(table_label=="results | Semi-final 1 [2]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 3: semi-final 2: songs, artist, overall points and placing - unnecessary 
results_2008_to_2026_filter %>% 
  # filter for table 
  filter(table_label=="results | Semi-final 2 [3]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 4: final: songs, artist, overall points and placing - unnecessary 
results_2008_to_2026_filter %>% 
  # filter for table 
  filter(table_label=="results | Final [4]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))

# table 5: 
results_2008_to_2026_filter %>% 
  # filter for table 
  filter(table_label=="results | Semi-final 1 [5]") %>% 
  # remove NA columns
  select(where(~!all(is.na(.))))
```

### Semi-final 1

```{r}

```

### Semi-final 2

```{r}

```

### Final 

```{r}

```


## Other 

```{r}
results_clean <- results %>%
  mutate(
    table_type = case_when(
      str_detect(table_label, "allocation draw") ~ "allocation_draw",
      str_detect(table_label, "Detailed voting results") ~ "detailed_voting",
      str_detect(table_label, "Final") ~ "final",
      str_detect(table_label, "Semi-final [12]") ~ "semifinal_summary",
      str_detect(table_label, "Semi-final \\d \\[") ~ "semifinal_detailed",
      str_detect(table_label, "^results \\| table") ~ "legacy_table",
      TRUE ~ "other"
    ),
    semi_final_number = case_when(
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  )

country_cols <- results_clean %>%
  select(where(is.numeric)) %>%
  select(-year, -r_o, -n, -no, -points, -number, 
         -starts_with("televote_"), -starts_with("jury_vote_"),
         -starts_with("combined_"), -starts_with("rest_"), 
         -semi_final_number) %>%
  colnames()

# ============================================================================
# EXTRACT VOTING DATA (LONG FORMAT) - FIXED
# ============================================================================

voting_data <- results_clean %>%
  select(
    year, table_label, table_type, country, 
    song, semi_final_number, artist, 
    all_of(country_cols)
  ) %>%
  pivot_longer(
    cols = all_of(country_cols),
    names_to = "country_voting",
    values_to = "votes_awarded"
  ) %>%
  filter(!is.na(votes_awarded), votes_awarded > 0, !is.na(country)) %>%
  mutate(
    country_voting = str_to_title(str_replace_all(country_voting, "_", " ")),
    country_performing = country
  ) %>%
  select(year, table_label, table_type, semi_final_number, 
         country_performing, song, country_voting, votes_awarded)

```

```{r}

head(voting_data, 20)
cat("\n✓ Voting data created successfully!\n")
cat("Rows:", nrow(voting_data), "\n")

# ============================================================================
# OUTPUT 1: SEMIFINAL TOTALS
# ============================================================================

eurovision_semifinal_totals <- voting_data %>%
  filter(table_type %in% c("semifinal_summary", "other")) %>%
  mutate(
    semi_final_number = case_when(
      !is.na(semi_final_number) ~ semi_final_number,
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(semi_final_number)) %>%
  group_by(year, semi_final_number, song, country_performing) %>%
  summarise(
    total_votes = sum(votes_awarded, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(country = country_performing) %>%
  select(year, semi_final_number, song, country, total_votes) %>%
  arrange(year, semi_final_number, desc(total_votes))

head(eurovision_semifinal_totals, 20)
cat("\nSemi-final totals rows:", nrow(eurovision_semifinal_totals), "\n")

# ============================================================================
# OUTPUT 2: FINAL TOTALS
# ============================================================================

eurovision_final_totals <- voting_data %>%
  filter(table_type == "final") %>%
  group_by(year, song, country_performing) %>%
  summarise(
    total_votes = sum(votes_awarded, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(country = country_performing) %>%
  select(year, song, country, total_votes) %>%
  arrange(year, desc(total_votes))

head(eurovision_final_totals, 20)
cat("\nFinal totals rows:", nrow(eurovision_final_totals), "\n")

# ============================================================================
# OUTPUT 3: SEMIFINAL VOTING BY COUNTRY (LIST)
# ============================================================================

eurovision_semifinal_voting_by_country <- voting_data %>%
  filter(table_type %in% c("semifinal_summary", "other")) %>%
  mutate(
    semi_final_number = case_when(
      !is.na(semi_final_number) ~ semi_final_number,
      str_detect(table_label, "Semi-final 1") ~ 1L,
      str_detect(table_label, "Semi-final 2") ~ 2L,
      str_detect(table_label, "Semi-final 3") ~ 3L,
      TRUE ~ NA_integer_
    )
  ) %>%
  filter(!is.na(semi_final_number)) %>%
  select(year, semi_final_number, song, country_performing, 
         country_voting, votes_awarded) %>%
  arrange(year, semi_final_number) %>%
  nest(data = -year) %>%
  deframe()

cat("\nSemi-final voting by country: list with", 
    length(eurovision_semifinal_voting_by_country), "years\n")
if (length(eurovision_semifinal_voting_by_country) > 0) {
  head(eurovision_semifinal_voting_by_country[[1]], 15)
}

# ============================================================================
# OUTPUT 4: FINAL VOTING BY COUNTRY (LIST)
# ============================================================================

eurovision_final_voting_by_country <- voting_data %>%
  filter(table_type == "final") %>%
  select(year, song, country_performing, country_voting, votes_awarded) %>%
  arrange(year) %>%
  nest(data = -year) %>%
  deframe()

cat("\nFinal voting by country: list with", 
    length(eurovision_final_voting_by_country), "years\n")
if (length(eurovision_final_voting_by_country) > 0) {
  head(eurovision_final_voting_by_country[[1]], 15)
}

# ============================================================================
# OUTPUT 5: SEMIFINAL ALLOCATION DRAW
# ============================================================================

eurovision_semifinal_allocation_draw <- results_clean %>%
  filter(table_type == "allocation_draw") %>%
  select(year, country, semi_final_assigned = semi_final_number, 
         broadcaster, language) %>%
  filter(!is.na(semi_final_assigned), !is.na(country)) %>%
  arrange(year, semi_final_assigned, country)

head(eurovision_semifinal_allocation_draw, 20)
cat("\nSemi-final allocation draw rows:", nrow(eurovision_semifinal_allocation_draw), "\n")

# ============================================================================
# DATA QUALITY CHECKS
# ============================================================================

cat("\n\n=== DATA QUALITY REPORT ===\n\n")

cat("📈 SUMMARY STATISTICS:\n")
cat("✓ Semi-final totals:", nrow(eurovision_semifinal_totals), "rows\n")
cat("✓ Final totals:", nrow(eurovision_final_totals), "rows\n")
cat("✓ Semi-final voting years:", length(eurovision_semifinal_voting_by_country), "\n")
cat("✓ Final voting years:", length(eurovision_final_voting_by_country), "\n")
cat("✓ Allocation draw entries:", nrow(eurovision_semifinal_allocation_draw), "\n")

```

# Inspect new df

```{r}
df_contest_format     <- read_csv(here("inst/extdata/eurovision_contest_format.csv"))
df_participants_info  <- read_csv(here("inst/extdata/eurovision_participants.csv"))
df_semifinal_results  <- read_csv(here("inst/extdata/eurovision_semifinal_results.csv"))
```

